\vspace{-1.8cm} \Large Demystifying Ambient Seismic Noise Processing: \\Automating source analysis and processing decisions in the presence of coherent anthropogenic, natural, and sensor noise \vspace{-0.5cm}

### 2017-02-17 paper.tex
finished first draft

\vspace{-1.8cm}
\subsection*{Motivation} 
\vspace{-0.2cm}
****motivation for near surface seismic imaging****
\par
Distributed acoustic sensing (DAS) probes a fiber optic cable (single or multi-mode \cite{Miller2016}) with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors. For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}. DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
\par
***motivation for using ambient noise****current issues in ambient noise*****
\par
There have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}. In fact, we can sacrifice some ground-to-sensor coupling for easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}. By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor are becoming a reality. When the interrogator unit is probing a section of fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes. 
\par
With large amounts of easily, cheaply recorded seismic data, we can only extract the full value of the data if we further automate the processing workflow.

\vspace{-0.3cm}
\subsection*{Significance} 
\vspace{-0.2cm}
This manuscript will walk through the basics (and some new tools) for processing ambient noise to do near-surface seismic imaging. Its purposes:
\vspace{-0.2cm}
\begin{itemize}
\item overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise
\vspace{-0.2cm}
\item show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved
\vspace{-0.2cm}
\item expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition continues to become easier and cheaper
\end{itemize}
\vspace{-0.2cm}
\par 
It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits undernear the Stanford University campus. This particular experiment is ideal because it demonstrates the wide variety of issues on the near horizon as we push for broader use of ambient noise. 
\par
We will discuss the noise level and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs (increasing data volumes): fiber loosely laying in existing telecommunications conduits. The array detects a wide variety of seismic noise sources that do not conform to the ideals of existing ambient noise theory: it sits in a seismically active region, 20 km from the Pacific ocean, 7 km from the San Francisco bay, with major highways on either side, a variety of roads with differing levels of traffic near the fiber, regular quarry blasts within 15 km, plumbing and HVAC systems throughout the site, multiple construction sites near the array, and foot and bicycle traffic throughout. With over 600 sensors continuously recording 50 samples per second since September 2016, manual inspection of most data is infeasible, making the automation tools shown in this paper critical to extracting subsurface information from these data.

@@

@@

@@

		\item New tool: Automatically detect \& filter persistent noise sources from short time spectra

@@

	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}. However, rather than using the first x amount of data, use a random sampling scheme over all the data so convergence rates get error bars.

@@

%\begin{abstract} 
%*****abstract words here - 150-200 Eileen*****
%\end{abstract}

%\section*{Motivation}
%****what's ambient noise used for? - 200-250 Eileen****

%****why use DAS? brief overview of its use****

%\section*{New Contributions}
%****what's new that we're doing, how it's different from what's been done before - 150 Eileen, little Fantine \& Yinbin****

%\section*{Background Theory}
%****ambient noise background - 150-220 Eileen****

%\section*{Experimental Overview}
%****overview of the array, potential noise sources, ambient noise field - 200 Eileen****
%****Figure : 2 panels: Installation + map****

%\section*{Data Processing}
%***overview of data processing workflow - 200-250 Eileen****
%***Figure : stft plots at a few different spots***
%***Mention whitening here***
%***Mention earthqukes and quarry blasts here***

%\subsection*{Estimating Sensor Coupling Response}
%***talk about different responses from different sensors and how to compensate for these, use many earthquakes for relative responses****
%***100 Eileen***
%***Figure : Boxed earthquake recording + response versus channel***

%% \subsection*{Localizing Persistent Noise Sources}
%% ***talk about localizing these things (motors, traffic, etc....) and show a little beamforming in small 1D parts of the array bandpassed to these persistent frequency peaks***

%% \subsection*{Removing Earthquakes and Quarry Blasts}
%% ****talk about getting rid of big events lighting up whole array****

%\subsection*{Removing Vehicle Noise}
%****talk about clustering and comparing noise sources****
%***500 Fantine \& Yinbin***
%***Figures : Before/after car mute + clusters and projection on map***
%***Little Eileen for sentence about why cars are a problem***

%% \subsection*{Removing Correlated Noise Sources}
%% *****talk about detecting temporally and spatially correlated noise sources, more generally than cars****

%% \subsection*{Traveltime Picking}
%% *** talk about traveltime picking before Rayleigh wave tomography****
%% *** STANDBY ***

%% \subsection*{Tomography}
%% **** talk about how near-surface tomography is set up****
%% *** STANDBY ***

%\section*{Results}
%***overview of results****
%*** Convergence rates
%***150 Eileen *** 
%***Figure : line plot of correlation coeeficients vs epoch/time***

%% \subsection*{Convergence Rate Improvement Due to Pre-Processing}
%***show comparisons of cross-correlations with/without sensor coupling response, and with scaling by energy per channel of cross-correlations (also show spectrum per channel of these)****
%\\
%******show comparisons of convergence rates with/without vehicles removed, and with vehicles just removed via STA/LTA jumps****

%% \subsection*{Tomography Along Array Lines}
%% ***show tomography results along multiple lines on campus****
%% *** STANDBY ***

%\section*{Summary}
%****summary words****
%***100 Eileen ***
\vspace{-0.4cm}
\subsection*{Biographies}
\small
\vspace{-0.2cm}
\textbf{Eileen R. Martin} (\texttt{ermartin@stanford.edu}) is a Ph.D. student in Stanford's Institute for Computational and Mathematical Engineering, and an M.S. student in the Geophysics Department. She first started working with ambient noise and DAS data from permafrost thaw monitoring experiments done by Lawrence Berkeley National Lab and the US Army Corps of Engineers.
\\
\textbf{Fantine Huot} (\texttt{fantine@sep.stanford.edu}) received her M.S. in Science and Engineering from the Ecole Normale Sup\'erieure des Mines de Paris in 2013. Following graduation, she worked as a software developer in France. In fall 2015, she started her Ph.D. in Geophysics at Stanford University and joined the Stanford Exploration Project. She works on addressing earth science problems by leveraging machine learning algorithms.
\\
\textbf{Yinbin Ma} \texttt{**email address} *****
\\
\textbf{Robert Cieplicki} (\texttt{robcie@stanford.edu}) is a PhD student in the Zoback Stress and Crustal Mechanics Research Group (Geophysics Department, Stanford University). He applies machine learning to horizontal wells drilled in shale hydrocarbon reservoirs to predict the efficiency of hydraulic fracturing.
\\
\textbf{Steve Cole} (\texttt{Steve.Cole@optasense.com}) is Manager of Integrated Analysis at OptaSense. He has worked in the oil and gas industry for over 30 years at companies including Fugro, Chevron, and 4th Wave Imaging, which he co-founded. He has a Ph.D. in geophysics from Stanford.
\\
\textbf{Martin Karrenbach} (\texttt{***email address}) *******
\\
\textbf{Biondo L. Biondi} (\texttt{**email address}) ****

@@

\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recognition}, vol. 38, no. 11, pp. 1857-1874, 2005.

>>>>

\vspace{-1.8cm}
\subsection*{Motivation} 
\vspace{-0.2cm}
By measuring the speed of seismic waves propagating in the Earth's near surface, we can image the top tens to hundreds of meters of the subsurface, with deeper features being resolved by lower frequencies. These seismic velocity images can be interpreted to evaluate earthquake or landslide risk, to detect permafrost, to find sinkholes or tunnels, or to track near surface changes related to drilling activities. Additionally, in cases of complex near-surface conditions, accurate near-surface images can be used to improve deeper seismic images.
\par
By cross-correlating noise recorded at any two receivers in an array, we can extract signals mimicking an active seismic survey with a source at any receiver, virtual source response functions \cite{Lobkis2001}, \cite{Lin2008}, \cite{Wapenaar2010A}, \cite{Wapenaar2010B}. Thus, when active sources are too costly or logistically prohibitive, passive seismic can be a good option for near surface imaging. However, the theory is limited by the assumption of homogeneous uncorrelated sources \cite{Wapenaar2010A}. Non-ideal sources can cause artifacts in extracted velocities, but with careful processing experts can overcome these limitations \cite{Bensen2007}, \cite{Daskalakis2016}, \cite{Zhan2013}, even for anthropogenic sources occurring at higher frequencies \cite{Girard2016}, \cite{Martin2015}, \cite{Martin2016}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. Acquisition is a further issue: nearly all of the past ambient noise studies on dense arrays have been temporary arrays which were labor-intensive to install and maintain \cite{Lin2008}, \cite{Nakata2011}, \cite{Nakata2015}, \cite{Zeng2017}. 
\par
Distributed acoustic sensing (DAS) probes a fiber optic cable with a laser interrogator unit (IU) to repurpose that fiber as a series of strain or strain rate sensors. For cost reasons, most experiments use straight fibers, but these fibers measure strain along the axis of the fiber, so some experimenters are now wrapping the fiber in a helix \cite{Kuvshinov2016}. DAS is being increasingly adopted in the energy industry for microseismic monitoring \cite{Webster2013} and time-lapse seismic surveys \cite{Daley2013}, \cite{Mateeva2014}, \cite{Bakku2015}, \cite{Miller2016} because it allows for a dense series of seismic sensors to communicate over long distances while they are run by a single power source, and those sensors can be permanently left in the same location.
\par
The maintenance cost of permanent node arrays motivated the use of buried DAS, and there have now been several ambient noise experiments where fiber optics were buried in shallow trenches so the fiber was directly coupled to the ground \cite{AjoFranklin2015}, \cite{Martin2015}, \cite{Martin2016}, \cite{Zeng2017}. In fact, we can sacrifice some ground-to-sensor coupling in favor of easier installation by running fibers in existing telecommunications conduits \cite{Martin2017}. By running fibers in existing conduits, or even plugging into existing dark fiber, on-demand repeatable seismic studies in urban areas requiring little manual labor will soon be a reality. When the interrogator unit probes a fiber, the data can serve many purposes: near surface imaging with ambient noise, earthquake detection and location (and local site response) \cite{Martin2017}, tracking vehicle traffic, potentially even detecting underground tunneling, studying the status of fluid pump systems nearby, or detecting intrusion into manholes. 
\par
With ambient noise data becoming increasingly easy to record, we can only extract its full value if we further automate the processing workflow.

\vspace{-0.5cm}
\subsection*{Significance} 
\vspace{-0.2cm}
This manuscript will walk through the basics and some new tools for processing ambient noise for near-surface seismic imaging. Its purposes: (1) overview of ambient noise processing for statisticians or mathematicians to encourage them to develop data analysis tools or theory for working with non-ideal ambient noise, (2) show optical engineers an application where an optical measurement device's noise level and sensitivity should be improved, and (3) expose gepohysicists to an emerging acquisition method and a few machine learning tools and metrics to quantify processing decisions, a task that will become increasingly more difficult as data acquisition becomes easier and cheaper.
\par 
It is hard to build intuition about amibent noise processing without looking at real data, so we will present this tutorial in the context of a case study: a figure-eight-shaped array of 2.4 km of fiber optics lying loosely in existing telecommunications conduits underneath the Stanford University campus. This particular experiment is ideal because it demonstrates the wide variety of issues on the near horizon as we push for broader use of ambient noise. 
\par
We will discuss the noise level and sensitivities particular to recording with DAS, and a novel deployment that can greatly reduce costs (increasing data volumes): fiber loosely laying in existing telecommunications conduits. The array detects a wide variety of seismic noise sources that do not conform to the ideals of existing ambient noise theory: it sits in a seismically active region, 20 km from the Pacific ocean, 7 km from the San Francisco bay, with major highways on either side, a variety of roads with differing levels of traffic near the fiber, regular quarry blasts within 15 km, plumbing and HVAC systems throughout the site, multiple construction sites near the array, and foot and bicycle traffic throughout. With over 600 sensors continuously recording 50 samples per second since September 2016, manual inspection of most data is infeasible, making the automation tools shown in this paper critical to extracting subsurface information from these data.

\vspace{-0.5cm}

@@

	\vspace{-0.8cm}

@@

	\vspace{-0.4cm}

@@

		\item New: Automatically detect \& filter persistent noise sources from short time spectra

@@

	\vspace{-0.5cm}
	\item Measuring improvements: to compare old and new processing methods measure convergence of virtual source response function estimates (both cross-coherence and cross-correlation) as more data is incorporated using the correlation coefficient like \cite{Seats2012}, but use a random sampling scheme over all data so convergence rates have error bars.

@@

\vspace{-0.9cm}
\subsection*{Biographies}
\small
\vspace{-0.2cm}
\textbf{Eileen R. Martin} (\texttt{ermartin@stanford.edu}) is a Ph.D. student in Stanford's Institute for Computational and Mathematical Engineering, and an M.S. student in the Geophysics Department. She is also affiliated with Lawrence Berkeley National Lab where she collaborates with a team on fiber optic monitoring of permafrost thaw and on the use of dark fiber for event detection.
\\
\textbf{Fantine Huot} (\texttt{fantine@sep.stanford.edu}) received her M.S. in Science and Engineering from the Ecole Normale Sup\'erieure des Mines de Paris in 2013. Following graduation, she worked as a software developer in France. In fall 2015, she started her Ph.D. in Geophysics at Stanford University and joined the Stanford Exploration Project. She leverages machine learning algorithms to address earth science problems.
\\
\textbf{Yinbin Ma} (\texttt{yinbin@stanford.edu}) is a Ph.D. student in the Institute for Computational and Mathematical Engineering at Stanford. He works on time-lapse seismic imaging at the Stanford Exploration Project.
\\
\textbf{Robert Cieplicki} (\texttt{robcie@stanford.edu}) is a PhD student in the Zoback Stress and Crustal Mechanics Research Group (Geophysics Department, Stanford University). He applies machine learning to horizontal wells drilled in shale hydrocarbon reservoirs to predict the efficiency of hydraulic fracturing.
\\
\textbf{Steve Cole} (\texttt{Steve.Cole@optasense.com}) is Manager of Integrated Analysis at OptaSense. He has worked in the oil and gas industry for over 30 years at companies including Fugro, Chevron, and 4th Wave Imaging, which he co-founded. He has a Ph.D. in geophysics from Stanford.
\\
\textbf{Martin Karrenbach} (\texttt{***email address}) *******
\\
\textbf{Biondo L. Biondi} (\texttt{**email address}) ****

@@

\bibitem{Liao2005} T.W. Liao, "Clustering of time series data- a survey," \textit{Pattern Recogn.}, vol. 38, no. 11, pp. 1857-1874, 2005.
